Lexicalized Domain Representations for Neural Machine Translation

### 2019-05-21 2019/emnlp/sparse.tex
start fixing part 2

\subsection{Problem statement \label{ssec:statement}}
\fyTodo{Use math for notation, label equations and such}
\begin{definition}{Domain}
\label{def:domain}
A domain is probabilistic distribution of translation $(x,y)$ in a given corpora. Given corpora $C$, we denote $D(C)$ the domain of $C$.
\end{definition}
\fyTodo{We need more: the source, the target, etc}
In machine translation, the loss function for training is:
\begin{equation}\label{eq:1}
\begin{split}
E_{(x,y) \sim p_{rel}}[-log(p_{\theta}(y|x))] &= \\
= E_{x \sim p_{rel}(x)}[KL(p_{rel}(y|x) & \mid p_{\theta}(y|x))] + \\
+ E_{x \sim p_{rel}} &[H(p_{rel}(y|x))]
\end{split}
\end{equation}.

@@

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc||c|}
\hline
Model & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
\hline
\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{8}{l}{English$\rightarrow$French} \\
\hline
$Mixed$               & 65.42$_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 &37.43 \\
\hline
$\mathtt{FT} (EMEA)$  & 72.06$_{47.33}$ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & 20 \\
$\mathtt{FT} (EPPS)$    & 35.47 $ $ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & 29.87 \\
$\mathtt{FT} (ECB)$     & 21.93 $ $ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & 22.23 \\
\hline
$DC$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & 37.24\\
 \hline
$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & 38.97 \\
\hline
$LDR_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & 38.57\\
$LDR^{v2}_{oracle}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & 32.6\\
$LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & 38.16\\
$LDR_{pred}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & 38.81\\
$LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 &  - & - & - & - \\
\hline
\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{8}{l}{English$\rightarrow$German} \\
\hline
$Mixed$               & 57.37 $_{37.94}$ & 23.1 & 63.54 & 17.39 & 11.51 & 14.29 & 31.2\\
\hline
$\mathtt{FT} (EMEA)$  & 65.64$_{44.71}$ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & 20.53\\
$\mathtt{FT} (EPPS)$   & 24.90 $ $ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & 18.89 \\
$\mathtt{FT} (ECB)$    & 41.8 $ $ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & 25.89\\
\hline
$DC$                      & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & 32.50\\
\hline
$\mathtt{WDCMT}$ & 64.05 $_{38.54}$ & 22.82 & 67.00 & 15.94 & 11.45 & 13.66 & 32.49\\
\hline
$LDR_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & 32.2\\
$LDR^{v2}_{oracle}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & 32.6\\
$LDR_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & 32.1\\
$LDR_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & 31.87\\
$LDR_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & - & - & - & avg \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the RNN systems\label{tab:results-rnn}}
\end{table*}
% \clearfloat

>>>>

\subsection{Multi-domain machine translation \label{ssec:statement}}

Multi-domain machine translation is formalized as follows: we assume that our observations correspond to  domain-tagged sentence pairs $[(x,y),i]$, with $x$ in the source language, $y$ in the target language and $i$ a domain tag in $[1\dots d]$. We further assume a two-stage sampling process: first select a domain $i$ according to $p(i)$, then select a sentence pair according to a domain specific distribution $D_i$. Our objective is to find a set of parameters $\{\theta_1 \dots \theta_d \} \in \mathbb{R}^D \times \mathbb{R}^D \dots \times \mathbb{R}^D$ minimizing:
\begin{equation} \label{eq:loss}
\begin{split}
\sum_{i \in [1..d]} p(i) E_{(x,y) \sim D_{i}} [-log(p_{\theta}(y|x,i))]
\end{split}
\end{equation}
The training data for domain $i$ is denoted $C_i$.

A straightforward solution is to process each domain separately, computing the value $\theta_i^*$ that minimizes the empirical loss on the training corpus $C_i$. This strategy is only effective if we have sufficient training data for each domain; when this is not the case, some estimates $\theta_i^*$ may be far from their optimal value. \fyDone{Risk}. 
The alternative we consider here constraints each parameter $\theta_i$ to be made of two parts: one ($\theta_s \in R^{D_g}$) is shared across all domains, while the second part $\theta_i \in \mathbb{R}^{D_i}$ is only used in domain $i$.
The parameter set is now much more constrained, yet we expect that tying parameters across domains will yield better estimates for $\theta_s$ due to a larger training corpus. In this setting, the optimization program defined by equation~\eqref{eq:loss} can no longer be performed separately for each training corpus.

%%% 
\fyTodo{Use math for notation, label equations and such}
\begin{definition}{Domain}
\label{def:domain}
A domain is probabilistic distribution of translation $(x,y)$ in a given corpora. Given corpora $C$, we denote $D(C)$ the domain of $C$.
\end{definition}
\fyTodo{We need more: the source, the target, etc}
In machine translation, the loss function for training is:
\begin{equation}\label{eq:1}
\begin{split}
E_{(x,y) \sim p_{rel}}[-log(p_{\theta}(y|x))] &= \\
= E_{x \sim p_{rel}(x)}[KL(p_{rel}(y|x) & \mid p_{\theta}(y|x))] + \\
+ E_{x \sim p_{rel}} &[H(p_{rel}(y|x))]
\end{split}
\end{equation}.

@@

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc||c|}
\hline
Model & EMEA & EPPS & ECB & Khresmoi & NewsTest2014 & IWSLT2010 & Avg. \\
\hline
\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{8}{l}{English$\rightarrow$French} \\
\hline
$Mixed$               & 65.42$_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 &37.43 \\
\hline
$\mathtt{FT} (EMEA)$  & 72.06$_{47.33}$ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & 20 \\
$\mathtt{FT} (EPPS)$    & 35.47 $ $ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & 29.87 \\
$\mathtt{FT} (ECB)$     & 21.93 $ $ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & 22.23 \\
\hline
$DC$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & 37.24\\
 \hline
$\mathtt{WDCMT}$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & 38.97 \\
\hline
$LDR_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & 38.57\\
$LDR^{v2}_{oracle}$   & 71.7$_{46.41}$ & 34.24& 52.37 & 29.21 & 22.7& 20.69 & 38.48\\
$LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & 38.16\\
$LDR_{pred}$        & 72.76 $_{46.35}$ & 35.1 & 50.38 & 30.02 & 23.28 & 21.32 & 38.81\\
$LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 &  - & - & - & - \\
\hline
\multicolumn{8}{l}{} \\[-9pt]
\multicolumn{8}{l}{English$\rightarrow$German} \\
\hline
$Mixed$               & 57.37 $_{37.94}$ & 23.1 & 63.54 & 17.39 & 11.51 & 14.29 & 31.2\\
\hline
$\mathtt{FT} (EMEA)$  & 65.64$_{44.71}$ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & 20.53\\
$\mathtt{FT} (EPPS)$   & 24.90 $ $ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & 18.89 \\
$\mathtt{FT} (ECB)$    & 41.8 $ $ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & 25.89\\
\hline
$DC$                      & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & 32.50\\
\hline
$\mathtt{WDCMT}$ & 64.05 $_{38.54}$ & 22.82 & 67.00 & 15.94 & 11.45 & 13.66 & 32.49\\
\hline
$LDR_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & 32.2\\
$LDR^{v2}_{oracle}$   & 63.27 $_{38.16}$ & 21.83 & 69.55 & 15.7 & 11.32 & 13.98 & 32.6\\
$LDR_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & 32.1\\
$LDR_{pred}$        & 63.17 $_{39.92}$ & 22.51 & 64 & 15.94 & 11.32 & 14.27 & 31.87\\
$LDR_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & - & - & - & avg \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the RNN systems\label{tab:results-rnn}}
\end{table*}
% \clearfloat
